Predicting Sequence Variability from Voronoi Tessellation of Proteins

### 2015-02-25 manuscript/voronoi/manuscript_voronoi.tex
minor changes. finally gitignore works!
Used this http://stackoverflow.com/questions/927358/how-to-undo-the-last-commit
to fix the gitignore problem globally.
minor mn modification

    Over the past decade, extensive attempt has been made to predict protein sequence evolution from structural properties.  A variety of site-specific structural characteristics have been proposed, including the Relative Solvent Accessibility (xx), Contact Number (xx), measures of thermodynamic stability changes due to mutations at individual sites in proteins \citep[e.g., the quantity ddG rate in][]{}(xx echave 2014), and measures of local flexibility, such as the Debye-Waller factor (hereafter B factor) (xx) or flexibility measures based elastic network models (xx Bahar et al.) and Molecular Dynamics (MD) simulations (xx shahmoradi 2014). \\

    Although structural characteristics have been individually extensively studied and explored with regards to their association with sequence evolution, it is yet unknown whether these seemingly independent quantities are merely different manifestations of a more fundamental underlying characteristics of individual sites in proteins or each influence the sequence evolution independently. It is perceivable that quantities such as B factor, RSA, and CN can all serve as {\it approximate} measures of local packing density of individual sites in proteins, or the local flexibility of individual amino acids. Recently, Huang et al 2014 (xx) argued, through an extensive mathematical model, for Contact Number as the primary determinant of sequence variability, as opposed to local flexibility. Nevertheless, site-specific flexibility is often represented by $C_\alpha$ atomic B factor, a quantity that is not necessarily a bais-free measure of the amino acid flexibility as a whole in a given site in protein. A more accurate measure of amino acid flexibility can be obtained from the accessible volume to each site in protein structure. An estimate of the accessible volume can be obtained through a quantity widely known as Contact Number (xx). For a given $C_\alpha$ backbone atom representing an amino acid in protein, the contact number is defined as the number of $C_\alpha$ atoms within a radius ($r$) of the reference $C_\alpha$ atom.  This definition of CN however, depends on an arbitrary parameter: the radius of neighborhood around the reference site.  An alternative more accurate definition of site-specific flexibility can be obtained from partitioning of the protein 3D structure. \\

>>>>

    A variety of site-specific structural characteristics have been proposed over the past decade to predict protein sequence evolution from structural properties. Among the most important and widely discussed are the Relative Solvent Accessibility (RSA) (xx), Contact Number (xx), measures of thermodynamic stability changes due to mutations at individual sites in proteins \citep[e.g., the quantity ddG rate in][]{}(xx echave 2014), and measures of local flexibility, such as the Debye-Waller factor (hereafter B factor) (xx) or flexibility measures based elastic network models (xx Bahar et al.) and Molecular Dynamics (MD) simulations (xx shahmoradi 2014). \\

    Although structural characteristics have been individually extensively studied and explored with regards to their association with sequence evolution, it is yet unknown whether these seemingly independent quantities are merely different manifestations of a more fundamental underlying characteristics of individual sites in proteins or each influence the sequence evolution independently. It is perceivable that quantities such as B factor, RSA, and CN, all serve as a proxy measures of local packing density of individual sites in proteins, or the local flexibility of individual amino acids. Recently, Huang et al 2014 (xx) argued, through an extensive mathematical model, for Contact Number as the dominant factor in sequence variability patterns, as opposed to local flexibility. It is notable that site-specific flexibility is often represented by $C_\alpha$ atomic B factor, a quantity that is not necessarily an unbiased measure of the amino acid flexibility as a whole in a given site in protein. A more accurate measure of amino acid flexibility can be obtained from the accessible volume to each site in protein structure. An estimate of the accessible volume can be obtained through a quantity widely known as Contact Number (xx). For a given $C_\alpha$ backbone atom representing an amino acid in protein, the contact number is defined as the number of $C_\alpha$ atoms within a radius ($r$) from the reference $C_\alpha$ atom.  This definition of CN however, depends on an arbitrary parameter: the radius of neighborhood around the reference site.  An alternative definition has been proposed in recent years, the Weighted Contact Number (WCN), which has been circumvents the free-parameter $r$ in the definition of CN by more accurate definition of site-specific flexibility can be obtained from partitioning of the protein 3D structure. \\
